\subsection[GPU Compute Model]{GPU Compute Model\footnote{by Anthony Gutierrez}}
\label{sec:gpu}

GPUs have become an important part of the system design for high-performance
computing, machine learning, and many other workloads. Thus, we have integrated
a compute-based GPU model into gem5~\cite{GutierrezBeckmann2018-amdAPU}. The
GPU model is based on the AMD Graphics Core Next (GCN) architecture
\footnote{http://amd.com/en/technologies/gcn} and currently
supports the AMD GCN3 ISA.

While the initial release of the GPU model supports the GCN architecture and
GCN3 ISA, the design provides enough flexibility to implement any
modern GPU architecture that follows a single-instruction, multiple-data
(SIMD) execution model. This is achieved by following the gem5
convention of hiding ISA- and architecture-specific functionality behind a
generic API. By following this convention, the GPU core models
can be made generic enough to be used across a variety of GPU ISAs, similar
to the CPU models.

\subsubsection[GPU Thread Context]{GPU Thread Context}
At a high level the GPU core's execution model can be described as a
single-instruction, multiple-thread (SIMT) execution model. In
this execution model threads (i.e., SIMD lanes) are grouped together into
a \textit{wavefront}
(using OpenCL terminology, also called a warp in CUDA terminology) and execute the
same instruction in lock-step on a single SIMD unit.

The wavefront class in gem5 includes all the relevant GPU thread state and
executes on a single SIMD unit within a compute unit (CU). Multiple wavefronts
can be combined to form a \textit{work-group} (also called
a thread block in CUDA terminology) and multiple work-groups can be combined to
form the GPU kernel. While wavefronts map to single SIMD units, a work-group maps
to a CU.

\subsubsection[Compute Unit Pipeline]{Compute Unit Pipeline}
The CU is the primary processing unit in the GPU, which contains many
CUs. The current model in gem5 supports in-order issue of SIMD and scalar
instructions processor.
The workhorse of the compute unit is the SIMD unit, which executes vector
ALU instructions. The CU model is highly configurable, allowing users to set the
number of SIMD units per CU as well as the width of each SIMD unit. In addition to
the SIMD units, each CU can have some arbitrary number of scalar units, scratch
pad memory, and global memory units (both vector and scalar memory).

The SIMD unit relies on large vector register files (VRF), however the GPU can
also be configured to include a scalar register file (SRF) for use with scalar
execution units. Additionally, vector ALU and memory instructions may contain
scalar operands, and thus vector operations require access to the SRF.

At a high level the CU pipeline modeled in gem5 consists of four stages: \textit{fetch},
\textit{scoreboard check}, \textit{schedule}, and \textit{execute}. The fetch stage
is responsible for arbitrating instruction fetch, storing fetched data,
and decoding instructions for all wavefronts on the CU. The scoreboard check stage is
responsible for ensuring that intra-wavefront dependences are
correctly handled before instructions can be scheduled. The schedule stage is
responsible for arbitrating among the wavefronts presented by
the scoreboard check stage for each execution unit type. Once an instruction is dispatched
to an execution unit it is guaranteed to execute.
This ``execute at execute'' approach is the same as used by the gem5 in-order and
out-of-order CPU models. The execute stage executes the instruction by calling the \textit{execute}
method on the instruction. It manages which resources are used and dictates the timing of
operations and register file reads and writes.

\subsubsection[Kernel Launch Flow]{Kernel Launch Flow}
Kernel launch is initiated by a host application making the appropriate kernel
launch call provided by the GPU's runtime library, for example the gem5 GPU model
currently supports the ROCm software stack for GPU compute. The kernel launch
call places a kernel launch packet (i.e., a descriptor of the kernel) in memory
and notifies the GPU hardware of
the kernel launch. The GPU contains a command processor (CP) that is responsible
for handling these calls from the host as well as scheduling kernels to the GPU.

When the host launches a kernel, the packet processor in gem5 receives a signal
indicating that a kernel launch packet is ready, extracts the packet from memory,
then reads important metadata from the packet in order to begin dispatching
work and initializing the kernel's state. After the launch packet has been parsed,
information about the kernel's resource requirements are sent to the dispatcher.

As previously described wavefronts are dispatched to a CU's SIMD units, while a work-group is mapped
to a single CU. Wavefronts and work-groups, however, are combined together to form an entire kernel
which is launched to the GPU. An entire work-group (and thus all wavefronts in a work-group) must be
launched together and on the same CU, however not all work-groups must be launched simultaneously.

Kernel launch is resource limited, and the dispatcher
in the GPU is responsible for tracking work-groups, both in-flight and waiting work-groups, and
dispatching them to the CUs as resources become available. In order for a work-group to be dispatched
the CU must have enough resources to support all of its wavefronts (e.g., register file space).
Once a work-group is ready for dispatch, the CP is notified and the CU resources are reserved.
Finally, the register file state for each wavefront in the work-group is initialized based
on the metadata extracted by the CP.

\subsubsection[Autonomous Data-Race-Free GPU Tester]{Autonomous Data-Race-Free GPU Tester\footnote{by Tuan Ta}}

The Ruby coherence protocol tester is designed for CPU-like memory systems that implement relatively strong memory consistency models (e.g., TSO) and hardware-based coherence protocols (e.g., MESI).
In such systems, once a processor sends a request to memory, the request appears globally to the rest of the system.
Without knowing implementation details of target memory systems, the tester can rely on the issuing order of reads and writes to determine the current state of shared memory.
However, existing GPU memory systems are often based on weaker consistency models (e.g., sequential consistency for data-race-free programs) and implement software-directed cache coherence protocols (e.g., the VIPER Ruby protocol which requires explicit cache flushes and invalidations from software to maintain cache coherence).
The order in which reads and writes appear globally can be different from the order they are issued from GPU cores.
Therefore, the previous CPU-centric Ruby tester is not applicable to testing GPU memory systems.

The gem5 simulator has added support for an autonomous random data-race-free testing framework to validate GPU memory systems.
The tester works by randomly generating and injecting sequences of data-race-free reads and writes that are synchronized by proper atomic operations and memory fences.
By maintaining the data-race freedom of all generated sequences, the tester is able to validate responses from the system under test.
The tester is also able to periodically check for forward progress of the system and report possible deadlock and livelock.
Once encountering a failure, the tester generates an event log that captures only memory transactions related to the failure, which significantly eases the debugging process.
Ta et al. show how the tester effectively detected bugs in the implementation of VIPER protocol in gem5~\cite{Ta2019gputesting}.
